---
title: "Several examples of Geometric Probability with R"
author: "Serafim Petrov"
date: "2023-10-21"
output:
  html_document: default
---

Geometric probability is a type of study that measures outcomes in terms of geometric proportions. It dates back to 18th century for the problem of Buffon's needle; and developed in Monte Carlo simulations-based approaches for problems such as integral estimation with the hit-and-miss method.

https://en.wikipedia.org/wiki/Buffon%27s_needle_problem

# Example 1:
It was an accident on a electricity line from 100 to 150-m line segment. What is the probability that the incident happened on the 120-130-m line segment?


```{r}
# Create a blank plot
plot(0, type = "n", xlim = c(90, 160), ylim = c(0, 0), xlab = "line length in meters", ylab = "")

lines(c(90, 160), c(0, 0), col = "grey", lwd = 5, lty = 'dashed')
lines(c(100, 150), c(0, 0), col = "black", lwd = 7)
lines(c(120, 130), c(0, 0), col = "red", lwd = 10)
points(c(120, 130), c(0, 0), col = "red", pch = 20)

```

The answer is: assuming a uniform distribution, the probability is the proportion of the length of the red line to the length of the black line: (130 - 120)/(150 - 100) = 1/5


# Example 2:
Pi estimation with the hit-and-miss method. Consider a circle with the radius = 1 inscribed in a square with a side = 2: the area of the square is 4; and the area of the circle is pi. Therefore, the proportion is area_circle/area_square = pi/4. 

Let x and y be independent and uniformly distributed random variables on the interval [-1, 1], that is each pair represents a coordinate inside of the square. Then the proportion of the coordinates inside of the circle x^2 + y^2 <= 1^2 to the all coordinates inside o the square is pi/4.

```{r}
set.seed(2024)
n <- 10000
circle <- (runif(n, -1, 1)^2 + runif(n, -1, 1)^2 <= 1)
pi.est <- mean(circle) * 4
replicate(10000, mean((runif(n)^2 + runif(n)^2 <= 1))*4) |> mean()
pi



# Create a blank plot
plot(c(-2, 2), c(-2, 2), type = "n", xlab = "", ylab = "", asp = 1)

# Plot the square
square <- cbind(c(-1, 1, 1, -1, -1), c(-1, -1, 1, 1, -1))
lines(square, col = "black")

# Plot the inscribed circle
theta <- seq(0, 2 * pi, length.out = 100)
circle_x <- cos(theta)
circle_y <- sin(theta)
circle_x <- circle_x * 1  # Scale the circle's x-coordinates by the radius
circle_y <- circle_y * 1  # Scale the circle's y-coordinates by the radius
lines(circle_x, circle_y, col = "red")

# Add coordinates
text(1.2, 1.2, "(1, 1)", col = "black")
text(-1.2, 1.2, "(-1, 1)", col = "black")
text(1.2, -1.2, "(1, -1)", col = "black")
text(-1.2, -1.2, "(-1, -1)", col = "black")

# Shade the circle
polygon(circle_x, circle_y, col = "pink", border = NA)


# Add X and Y coordinate axes
abline(v = 0, col = "black")  # Vertical line at X = 0
abline(h = 0, col = "black")  # Horizontal line at Y = 0
```


<!-- # Example 3: -->

<!-- x + y < 2 -->

<!-- ```{r} -->
<!-- # Create a blank plot -->
<!-- plot(c(0, 5), c(0, 5), type = "n", xlab = "", ylab = "", asp = 1) -->

<!-- # Define the coordinates of the square -->
<!-- square <- rbind(c(0, 0), c(0, 5), c(5, 5), c(5, 0), c(0, 0)) -->

<!-- # Plot the square -->
<!-- lines(square, col = "black") -->

<!-- # Add the line x*y < 2 -->
<!-- x <- seq(0, 10, length.out = 100) -->
<!-- y <- 2-x -->
<!-- y[y == Inf] = 100 -->
<!-- lines(x, y, col = "red") -->
<!-- ``` -->


# Example 3:
xy < 2

```{r}
mc_integral <- function(ftn, a, b, n){
  u = runif(n, a, b)
  x = sapply(u, ftn)
  return(mean(x)*(b-a))
}

f1 <- function(x) 2/x
a1 <- 2/5
b1 <-  5
n1 <- 100000

res_A <- a1*b1
res_B <- mc_integral(f1, a1, b1, n1)
res_C <- b1*b1 - (res_B + res_A)

res_C/(b1*b1) # xy > 2


# Create a blank plot
plot(c(0, 5), c(0, 5), type = "n", xlab = "X", ylab = "Y", asp = 1)

# Define the coordinates of the square
square <- rbind(c(0, 0), c(0, 5), c(5, 5), c(5, 0), c(0, 0))

# Plot the square
lines(square, col = "black")

# Add the line x*y < 2
x <- seq(0, 10, length.out = 100)
y <- 2 / x
y[y == Inf] = 100
lines(x, y, col = "red")
abline(v = 2/5, lty = "dashed", col = "blue")



```

$$
\int_{.4}^{5} 2/x dx= 2 ln(|x|) |_{.4}^{5} = 2(ln(5) - ln(.4)) \approx 5.051457

$$
$$
\int_{.4}^{5} \int_{2/x}^{5} dxdy = \int_{.4}^{5} (5-2/x) dx= (5x - 2ln(|x|))|_{.4}^{5} = 
25 - 2ln(5) - 2 +2ln(.4) \approx 17.94854
$$


# Example 4:
meeting problem: 


```{r}
wait1 = 30
wait2 = 30
n = 1000

set.seed(2024)
start_time <- Sys.time()
ex3 <- function(waiting1 = 30) {
  x = runif(1, 0, 90)
  y = runif(1, 30, 120)
  
  condition1 = x < y & x + waiting1 > y
  condition2 = y < x & y + waiting1 > x
  
  return(condition1 || condition2)
}
replicate(n, mean(replicate(n, ex3(waiting1 = wait1)))) |> mean()
print(Sys.time() - start_time)


```


$$
A = \int_{0}^{90} \int_{y=x+30}^{y=120} dxdy= \int_{0}^{90} (120 -x-30)dx = -(x^2/2 +90)|_{0}^{90} = 4050
$$

$$
B = \int_{60}^{90}\int_{y=30}^{y=x-30}dxdy = \int_{60}^{90} (x-60) dx = (x^2/2 - 60x)|_{60}^{90} = 450 
$$
$$
C = T - A - B = 8100 - 4050 - 450 = 3600
$$
Find how long to wait to reach the prob of 80%

```{r}
set.seed(2024)
for (i in 20:60) {print(c(i, mean(replicate(n,ex3(i)))))}
set.seed(2024)
sapply(20:60,function(x) mean(replicate(n,ex3(x))))
```


Arrive in exact time?


```{r}
set.seed(2023)
ex1 <- function(n){
  x = as.integer(runif(n, 0, 90))
  y = as.integer(runif(n, 30, 120))
  sum(x == y)/n 
}

replicate(10000, ex1(10000)) |> mean()

```











# Example 5:
3-dimensional problem:

```{r}
# Install and load the required packages
library(plotly)

# Define the radius of the sphere
radius <- 1

# Generate data points for the surface of the sphere
n_points <- 100
theta <- seq(0, 2 * pi, length.out = n_points)
phi <- seq(0, pi, length.out = n_points)
x_sphere <- radius * outer(cos(theta), sin(phi))
y_sphere <- radius * outer(sin(theta), sin(phi))
z_sphere <- radius * outer(rep(1, n_points), cos(phi))

# Define the coordinates of the cube vertices
cube_vertices <- matrix(c(
  -1, -1, -1,
  -1, -1, 1,
  -1, 1, -1,
  -1, 1, 1,
  1, -1, -1,
  1, -1, 1,
  1, 1, -1,
  1, 1, 1
), ncol = 3, byrow = TRUE)

# Create a 3D scatter plot using Plotly
plot_ly() %>%
  add_trace(
    x = cube_vertices[c(1, 2, 4, 3, 1), 1],
    y = cube_vertices[c(1, 2, 4, 3, 1), 2],
    z = cube_vertices[c(1, 2, 4, 3, 1), 3],
    type = "scatter3d", mode = "lines",
    line = list(color = "black", width = 2)
  ) %>%
  add_trace(
    x = c(x_sphere),
    y = c(y_sphere),
    z = c(z_sphere),
    type = "scatter3d", mode = "markers",
    marker = list(color = "red", size = 2),
    name = "Sphere"
  ) %>%
  layout(
    scene = list(aspectmode = "cube"),
    title = "Sphere Inside Cube"
  )
```


